---
title: 'Emergency 911 Calls'
output: html_document
---

***


```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(ggmap)
library(viridis)
library(plotly)
```


```{r echo=FALSE, warning=FALSE}

mydata <- read.csv("C:/Users/SergeyPC/Desktop/montcoalert/911.csv")

dim(mydata)

str(mydata)

summary(mydata)

```


```{r echo=FALSE, warning=FALSE}

mydata$zip <- factor(mydata$zip)

mydata <- mydata[,-9]

mydata$timeStamp <- as.POSIXct(mydata$timeStamp)

mydata$Date <- as.Date(mydata$timeStamp)

mydata <- separate(mydata, col = title, into = c("Type", "SubType"), sep = ":")

mydata$SubType <- gsub(" -", "", mydata$SubType)

```


### Creating new variables

```{r echo=FALSE, warning=FALSE}

mydata$Year <- year(mydata$timeStamp)
mydata$Month <- month(mydata$timeStamp)
mydata$Day <- day(mydata$timeStamp)
mydata$Hour <- hour(mydata$timeStamp)
mydata$Weekday <- weekdays(mydata$timeStamp)

mydata$Year <- factor(mydata$Year)
mydata$Month <- factor(mydata$Month)
mydata$Day <- factor(mydata$Day)
mydata$Hour <- factor(mydata$Hour)
mydata$Weekday <- factor(mydata$Weekday)

mydata <- mydata[,-7]

```

### Number of calls over the period of time and by type

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

by_date <- mydata %>% group_by(Date) %>% summarise(Total = n())

head(by_date)

ggplot(by_date, aes(Date, Total)) + geom_line(color = "blue", size = 1)+xlab("Месяц")+ylab("Количество")

by_date_type <- mydata %>% group_by(Date, Type) %>% summarise(Total = n())

by_date_type$Type <- factor(by_date_type$Type)

ggplot(by_date_type, aes(Date, Total)) + geom_line( aes(color = Type), size = 0.6)+xlab("Дата")+ylab("Количество")
```


```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

table(mydata$Year)

table(mydata[mydata$Year]$Month)

table(mydata[mydata$Year]$Month)

ggplot(mydata[mydata$Year==2016,], aes(Month, fill = Month)) + geom_bar() + theme(legend.position = "none") + ggtitle("Вызовы в службу 911 в 2016г по месяцам ")+xlab("Номер месяца")+ylab("Количество")

ggplot(mydata, aes(Hour, fill = Hour)) + geom_bar() + theme(legend.position = "none") + ggtitle("Вызовы в службу 911 в 2016г по часам")+xlab("Часы")+ylab("Количество")

ggplot(mydata, aes(Weekday, fill = Weekday)) + geom_bar() + theme(legend.position = "none") + ggtitle("Вызовы в службу 911 в 2016г по дням недели")+xlab("Дни недели")+ylab("Количество")

```

### How many calls based on type

```{r echo=FALSE, warning=FALSE, fig.width=8, fig.height=5}

table(mydata$Type)
prop.table(table(mydata$Type))


```

### How many calls from each subtypes in overall dataset and also subtypes within each type

```{r echo=FALSE, warning=FALSE, fig.width=10, fig.height=5.5}

top_subtypes <- as.data.frame(table(mydata$SubType))
top_subtypes <- top_subtypes[order(-top_subtypes$Freq),]
top10_subtypes <- top_subtypes[1:10,]
top10_subtypes$Perc <- top10_subtypes$Freq/sum(top_subtypes$Freq) * 100
top10_subtypes

gettop10subtypes <- function(type) {
  mytype <- subset(mydata, Type == type)
  mytype$SubType <- factor(mytype$SubType)
  mytype_subtypes <- as.data.frame(table(mytype$SubType))
  mytype_subtypes <- mytype_subtypes[order(-mytype_subtypes$Freq),]
  top10_types_substype <- mytype_subtypes[1:10,]
  top10_types_substype$Perc <- top10_types_substype$Freq/sum(mytype_subtypes$Freq) * 100
  return(top10_types_substype)
}

gettop10subtypes("EMS")

gettop10subtypes("Fire")

ggplot(gettop10subtypes("Fire"), aes(reorder(Var1, Freq), Freq, fill = Var1)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("Звонки в службу 911 с типом 'огонь'") + xlab("Подтип") + ylab("Количество звонков")  + coord_flip()

```

### Which zip codes have the highest number of calls

```{r echo=FALSE, warning=FALSE,fig.width=9, fig.height=5.5}

top_zip <- as.data.frame(table(mydata$zip))
top_zip <- top_zip[order(-top_zip$Freq),]
top10_zip <- top_zip[1:10,]

names(top10_zip) <- c("Zip", "Total")
top10_zip$Perc <- top10_zip$Total/sum(top_zip$Freq) * 100
top10_zip$Zip <- factor(top10_zip$Zip)

top10_zip

ggplot(top10_zip, aes(reorder(Zip, -Total), Total, fill = Zip)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("Топ 10 zip-кодов") + xlab("Zip-код (индекс) откуда поступало наибольшее количество звонков")+ylab("Количество")

```


### Which township contributes to majority of the calls


```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

length(unique(mydata$twp))

top_twp <- as.data.frame(table(mydata$twp))
top_twp <- top_twp[order(-top_twp$Freq),]
top10_twp <- top_twp[1:10,]

names(top10_twp) <- c("Twp", "Total")
top10_twp$Perc <- top10_twp$Total/sum(top_twp$Freq) * 100
top10_twp$Twp <- factor(top10_twp$Twp)

top10_twp

ggplot(top10_twp, aes(reorder(Twp, -Total), Total, fill = Twp)) + geom_bar(stat = "identity") + theme(legend.position = "none") + ggtitle("Топ городов с наибольшим количеством звонков") + xlab("Города с наибольшим количеством звонков") + ylab ("Количество")  + theme(axis.text.x=element_text(angle=45,hjust=0.5,vjust=0.5)) 

```


```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}


```

### How does the type vary among top 10 township

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

ggplot(mydata[mydata$twp %in% top10_twp$Twp, ], aes(twp, fill = Type )) + geom_bar(position = "dodge")  + theme(axis.text.x=element_text(angle=45,hjust=0.5,vjust=0.5))  + xlab("Города") + ggtitle("Топ 10 городов по звонкам в службу 911 с типом")+ylab("Количество")

```

### How does the Top 5 subtypes vary among top 10 twp

```{r echo=FALSE, warning=FALSE, fig.width=9, fig.height=5.5}

top3 <- top10_subtypes[1:5,]$Var1

sample <- mydata[mydata$twp %in% top10_twp$Twp,]
sample <- sample[sample$SubType %in% top3, ]
dim(sample)

ggplot(sample, aes(twp, fill = SubType )) + geom_bar(position = "dodge")  + theme(axis.text.x=element_text(angle=45,hjust=0.5,vjust=0.5))  + xlab("Города") + ylab("Количество") + ggtitle("Подтипы звонков совершенных в службу 911 в топ 10 городов")

```

